### 2013-05-24 bylaws.tex

% BEFORE CHANGES ARE MADE TO THIS DOCUMENT:
% -References will be automatically updated if any part is added, deleted, etc. 
%  However, if a sub part is moved to a different part, its references must be 
%  changed.
% -This document must be ratified by the House (as per the Constitution), 
%  then printed, signed, notarized, and placed in the House filing cabinet
%  if changes are to be officialized.

\documentclass{article}
% The xr package allows external references
\usepackage{xr-hyper}
\usepackage{hyperref}

% This package is useful for debugging label problems
% Comment out in final revision
%\usepackage{showkeys}

% Define the external document to be constitution for cross referencing purposes
\externaldocument{constitution}[https://github.com/ComputerScienceHouse/Constitution/blob/master/constitution.pdf?raw=true]

% Title page information
\title{Computer Science House By-Laws}
\author{Computer Science House Constitution Committee}
% Last Modified Date
\newcommand{\datechanged}{Last Updated \today}
\date{\datechanged}

% Fix margins
\setlength{\evensidemargin}{0in}
\setlength{\oddsidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{0in}
\setlength{\textheight}{8.5in}

% Use \article for articles and \asection for sections of articles.
% Automatically provide labels with the same article or section title.
\newcommand{\bylaw}[1]{\section{#1} \label{#1}} 
\newcommand{\bsection}[1]{\subsection{#1} \label{#1}}
\newcommand{\bsubsection}[1]{\subsubsection{#1} \label{#1}}
\renewcommand{\thesection}{By-Law \Roman{section}}
\renewcommand{\thesubsection}{Section \arabic{section}.\Alph{subsection}}
\renewcommand{\thesubsubsection}{\arabic{section}.\Alph{subsection}.\arabic{subsubsection}}

% Adding a \bsubsubsection -- I feel dirty
\setcounter{secnumdepth}{5}
\newcommand{\bsubsubsection}[1]{\paragraph{#1} \label{#1}}
\renewcommand{\theparagraph}{\arabic{section}.\Alph{subsection}.\arabic{subsubsection}.\Alph{paragraph}}

% Headings
\pagestyle{myheadings}
\markright{{\rm CSH Constitution \hfill \datechanged \hfill Page }}
% Reference example:
%Test reference \ref{House Objectives} House Objectives.

\begin{document}
% Title
\maketitle

% BY-LAW I - INTRODUCTION
\bylaw{Introduction}
By-Laws will expand upon, while not redefining, Articles laid out in the Constitution. By-Laws provide additional depth and definition to the processes and operations of House.

% BY-LAW II - GENERAL OPERATIONS OF THE HOUSE
\bylaw{General Operations of the House}
\bsection{Standard Operating Session}
The Standard Operating Session for Computer Science House is during the fifteen weeks of Fall and Spring semesters. The Summer Sessions, Intersessions, Institute breaks, and all end of Semester Exam Weeks are considered non-standard operating sessions. Unless explicitly stated otherwise, the requirements and expectations deﬁned in the constitution are for the Standard Operating Session.

% BY-LAW III - OPERATIONS OF THE FINANCIAL DIRECTORSHIP
\bylaw{Operations of the Financial Directorship}
\bsection{Amount of House Dues}
The amount of dues for Active Members living on floor will be fifty-five dollars (\$55.00) per Academic Semester. The amount of dues for Active Members living off floor twill be fifty dollars (\$50.00) per Academic Semester.
\bsection{Collection of House Dues}
At the start of the academic year, dues for on-floor members (for both semesters) are collected through the member’s RIT bill. Dues for off-floor members are to be collected at the start of each semester by the Financial Director. For any member who moves on or off-floor during the year, any difference between dues paid and dues owed should be collected.
\\*\\*
If a member is unable to pay dues upon request, they may appeal their situation to the Financial Director. If the Financial Director deems their situation appropriate, they may grant an extension for payment. If the Financial Director deems their situation inappropriate for extension, the member may appeal to the Executive Board. If the Executive board disagrees with the Financial Director, the Executive board may grant an extension. If the Executive Board agrees with the Financial director, the member’s payment is considered delinquent and the member’s privileges are revoked until payment can be made. This process may be repeated indefinitely at the discretion of the Financial Director and the Executive Board.
\bsection{Breakdown of Dues for Directorship Budgets}
\begin{center}
\begin{tabular}[c]{l c}
Directorship Name & Percentage of Dues \\
\hline
\hline
Operational Communications & 20\% \\
\hline
Evaluations \& Selections & 5\% \\
\hline
House History & 10\% \\
\hline
House Improvements & 15\% \\
\hline
Research and Development & 20\% \\
\hline
Social & 20\% \\
\hline
Accumulated & 10\% \\
\hline
\end{tabular}
\end{center}

The suggested total operational budget is \$5544. This figure is based on yearly estimated on-floor member dues (\$110 x 56 members = \$6160) minus the 10\% reserved for Accumulated (Accum). Money allocated for Accum and off-floor member dues are deposited into the CSH Account and saved.

\bsection{Expenditure Approval}
All House expenditures must be approved by both the Financial Director and the appropriate Director (from whose budget the funds will be drawn). Single expenditures may not exceed seventy-five dollars (\$75) and total expenditures may not exceed one-hundred dollars (\$100). A total expenditure is defined as all the funds drawn for a specific event, project, piece of equipment, or service.
\\* \\*
If the above amounts are to be exceeded, the expenditure must be approved by a Simple Majority vote at House meeting before the funds may be appropriated, as defined in \ref{Simple Majority}. It is required that, for expenditures exceeding \$300 in total funding, a Complete Project Proposal must be presented to House at a House meeting. This proposal includes the project budget, inventory of required resources, and a timeline for completion. It is also recommended that all expenditure approvals done at House meeting, regardless of amount, provide such a proposal.
\\* \\*
If an appropriate Directorship cannot be determined for an expenditure, it is to be brought up for approval at House meeting as a Miscellaneous expenditure and approved by a simple majority vote, regardless of the amount. Miscellaneous purchases are deducted directly from the CSH Account, and therefore must have final approval of the Financial Director.
\\* \\*
Funds allocated for a project not spent by the end of the Standard Operating Session \ref{Standard Operating Session} in which they were approved are voided, unless otherwise specified during the original voting process.

% BY-LAW IV - OPERATIONS OF THE EVALUATIONS DIRECTORSHIP
\bylaw{Operations of the Evaluations Directorship}
\bsection{The Introductory Process}
The Introductory Process is designed to provide an easy means for introductory members to meet existing House members, learn House history, demonstrate their involvement potential to House, and allow existing House members to evaluate them for Active Membership.
\bsubsection{The Evaluation Period}
The Process lasts between six (6) and ten (10) weeks, starting when the Evaluations Director initiates it. The Executive Board may hold a closed ballot Simple Majority vote to extend the Introductory Process.
\bsubsection{The Introductory Project}
Introductory members are expected to demonstrate initiative and an ability to work well in a group environment through the Introductory Project.
\bsubsection{Introductory Executive Board}
An Introductory Executive Board is elected by the introductory members to lead and manage the project, represent the introductory project participants to House, and evaluate participants based on their contribution to the project. The Introductory Executive Board consists of three members:
\begin{itemize}
	\item A President, who leads the project and acts as primary representative to House through weekly status reports at House Meeting.
	\item A Vice President, who assists the President in leadership and representation.
	\item A Treasurer, who manages the finances for the project.
\end{itemize}
Introductory Members also elect a Secretary to take notes, and any number of additional non-voting leaders may be elected to take on project-specific tasks.
\bsubsubsection{The Project} 
The Project is an annual fundraising event that takes place in the first semester of the Introductory Process. All of the proceeds go to a charity chosen in advance by the participants. The Executive Board is encouraged to discuss the Project and following an open ballot Simple Majority Vote, may choose to intervene in the Introductory Project if the Project does not appear to fulfill requirements.
\bsubsubsection{Introductory Project Advisors} 
The Evaluations Director may choose between one (1) and three (3) House members to be Introductory Project Advisors. The advisors assist in organizing the participants at the beginning of the Introductory Project and thereafter act as a resource to keep the Project running smoothly.
\bsubsection{The Introductory Packet}
Each Introductory Process participant, after the first week, is given two (2) weeks to return a document that should contain the following:
\begin{itemize}
	\item A signature from: all Resident members, each Executive Board member, all Introductory members who reside on Floor, and fifteen (15) Non-Resident and/or Alumni members (not including Non-Resident Executive Board members)
	\item A description of each Executive board position, including:
	\begin{itemize}
		\item The responsibilities of the position
		\item The name(s) of the position’s member(s)
		\item The position’s weekly meeting time
	\end{itemize}
	\item A list of seven (7) annual House social events
	\item A list of seven (7) major House technical achievements
\end{itemize}
\bsubsection{Expectations of an Introductory Member}
Before the end of the Process, an Introductory Member is expected to:
\begin{itemize}
\item Attend all House meetings during the process
\item Complete the Introductory Packet
\item Be an active participant in an the Introductory Project
\item Attend at least one House directorship meeting for each week of the process
\item Attend at least one House social event during the process
\item Attend at least two seminars relating to computing or electronics
\item Sign a copy of the Membership Agreement describing a House Member’s responsibilities and expectations.
\end{itemize}

\bsubsection{Introductory Evaluation}
The Introductory Evaluation is the process by which House chooses which Introductory Members to extend an offer of full Resident or Non-Resident Membership. Occurs once per academic year as part of the Introductory Process.
\bsubsubsection{Voting} 
On a member by member basis, this evaluation process determines if each Introductory member has successfully completed the Introductory Process requirements \ref{Expectations of an Introductory Member}. A Simple Majority vote with a quorum of two-thirds of all Active members is required for the evaluation to be valid. Neither absentee nor proxy votes are allowed. House may choose any of the Outcomes for each member.
\bsubsubsection{Outcomes} 
\renewcommand{\theenumi}{\alph{enumi}} % For this section, we want items to use letters
\begin{enumerate}
	\item Introductory Members may be offered Active Membership (\ref{Active Membership}) provided they meet the requirements described in \ref{Expectations of an Introductory Member}, at the discretion of House.
	\item If an Introductory Member fails to meet the requirements, his membership will be revoked. In addition, the participant is asked to find alternative housing as soon as possible in accordance with all applicable Residence Life policies regarding room changes.
	\item An Introductory member may be given a conditional to complete as a means of making up for missing requirements. A conditional may be proposed by any member present at the Introductory Evaluation and, if it is approved by the Evaluations Director, is then voted on by House. Each conditional consists of a set of additional requirements a deadline for completing them. When the deadline expires, the conditional will be brought before the Executive Board who will assess its completeness. A conditional member who does not meet these additional requirements forfeits membership.
\end{enumerate}
\bsection{Selection Processes}
\renewcommand{\theenumi}{\alph{enumi}} % For this section, we want items to use letters
\bsubsection{Selection Process for Students Attending RIT for at Least One Semester}
\begin{enumerate}
	\item The applicant submits an application to the Evaluations Director for review.
	\item The applicant participates in an informal interview with exactly three current Active, Alumni, or Honorary Members.
	\item The application materials are then reviewed at an Evaluations meeting and the members present vote on whether to accept the person as an introductory member.
\end{enumerate}
Note: Most membership privileges do not initiate until successful completion of the introductory process. This means that until the member has passed the introductory evaluation, the member does NOT have the right to vote on House issues and does NOT count towards quorum. The member does, however, have the right to use Computer Science House's facilities. The member must also pay dues.
\\* \\*
Additional Note: No hazing shall occur at any time during the Selection Process in accordance with the New York State Hazing Laws.
\bsubsection{Selection Process for First Semester and Entering RIT Students}
\begin{enumerate}
	\item During Spring semester, the Evaluations Director selects a group of Active Members to form a Selections Committee. The Selections Committee reviews applications and conducts interviews in accordance with the process prescribed by Residence Life.
	\item A subset of applications will be selected to move onto the House Fall Semester, are granted full membership privileges (including On-Floor status), and must participate in the Introductory Process as defined in \ref{The Introductory Process}. An additional subset will be given Off-Floor status, but otherwise receive the same privileges and responsibilities.
\end{enumerate}
Note: Most membership privileges do not initiate until successful completion of the introductory process. This means that until the member has passed the introductory evaluation, the member does NOT have the right to vote on House issues and does NOT count towards quorum. The member does, however, have the right to use Computer Science House's facilities. The member must also pay dues.
\\* \\*
Additional Note: No hazing shall occur at any time during the Selection Process in accordance with the New York State Hazing Laws.
\bsubsection{Selection Process for Honorary Members}
\begin{enumerate}
	\item A House Member submits to the Evaluations Director a nomination for a person they feel is deserving of Honorary Membership.
	\item The Evaluations Director performs some preliminary research on the candidate and presents the findings.
	\item The Executive Board decides whether or not to present the nomination to the House for a secret ballot House vote. If the Executive Board decides not to present the nomination to the House, the Selection Process ends and the candidate does not become an Honorary Member.
	\item The nomination is presented at a House Meeting for discussion and a Two-Thirds vote, as described in \ref{Two-Thirds}. Ballots are distributed and voting must remain open for a minimum of forty-eight hour period.
	\item The candidate is notified of their selection as an Honorary Member and presented with the honor.
\end{enumerate}
\bsubsection{Selection Process for Advisory Members}
\begin{enumerate}
	\item When there is a perceived need, the Executive Board may open nominations for Advisory Members. After an announcement at House Meeting, during a 72-hour period, any House Member may submit a nomination.
	\item After the close of the nomination collection period, the Executive Board will arrange some means for the House to meet with the nominees.
	\item A discussion of the candidates will be held at the following House meeting.
	\item Ballots will then be distributed for a secret ballot, fifty-percent House vote, as defined in \ref{Fifty Percent}. If more than two advisors are selected via this process, the Executive Board may call for a revote amongst the candidates selected in order to reduce the number of advisors to a workable level. If this decision is made, the candidates selected in the first ballot become the nominees for the second ballot and Step Three and Four are repeated. This recast process may not be repeated a second time.
	\item All candidates selected are notified of their acceptance as House Advisors and asked to accept or decline the selection. If one or more selected candidates decline the position, the Executive Board may begin a selection process for additional advisors. Note that selected candidates who accept their selection remain selected, this additional selection process is only a means to bring the number of advisors up to a workable level. If the additional advisor selection process is initiated, repeat Steps One through Five with the original nominations plus any additional nominations.
\end{enumerate}

\bsection{Evaluations Processes}
During the academic year, a House member is evaluated by two different processes. The first process, known as the Housing Evaluation, is an informal evaluation designed to recognize an individual's contributions to the House, as well as checkpoint an individual's involvement with the House. The second process, known as the Membership Evaluation, is responsible for determining those individuals who will have the option to continue Active Membership in the following year. These processes are separate from each other and should not interfere with one another. Semi-Annual Evaluations Meetings are open only to all current Active Members. All Executive Board members are expected to attend the Semi-Annual Evaluations Meetings to assist in the evaluations of House members.
\\* \\*
At the beginning of any Evaluation Process listed herein, the Evaluations Director must read the sections of the Computer Science House Constitution and By-Laws used during the respective Evaluation Process. It is incumbent upon each House member to provide the Evaluation Director with whatever information he or she feels is necessary to ensure an accurate evaluation.
\bsubsection{Housing Evaluation}
The Housing Evaluation process occurs during the academic year. It is performed late Fall Semester after the Introductory Evaluation and is led by the Evaluations Director. The Housing Evaluation is used to recognize an individual’s contributions to the House and distribute Housing Points (see \ref{Housing Priority System}).
\\* \\*
All Active members who wish to be evaluated are required to submit an activity report in which they may log projects worked on, social events and seminars hosted or attended, and any other information they deem relevant. A member’s failure to submit the activity report will result in that member receiving no housing points. Attendance is also required for a member to be evaluated, else they will receive no housing points. Any member unable to attend because of outstanding circumstances (e.g. away on co-op) may also submit an activity report and be voted on, though they may not vote. Members in this situation should contact the Evaluations Director at least one day prior to the evaluation.
\\* \\*
The participating members, on a member by member basis, review the activity report provided by each member. The participating members may then choose to award one, two, or three housing points to each member. Points are awarded on the basis of contributions to House. Above average members receive three points, average members two points, and below average members one point.
\\* \\*
A quorum of two-thirds of all Active (excluding Introductory Process Participants) Members is required for the evaluation to be valid. Neither Absentee nor Proxy votes are allowed.
\\* \\*
All votes during the Housing Evaluation Process must be Closed Ballot, Three Tiered Vote as defined in \ref{Three Tiered}.
\bsubsection{Membership Evaluation}
The Membership Evaluation process occurs once per academic year. It is performed as part of the Evaluation Process that takes place during the spring semester to comply with RIT Housing deadlines.
\bsubsubsection{Voting} 
A quorum of members, on a member by member basis, ensures that each Active member has met the minimum requirements in order to continue as an Active member in the following year. All Active Members will be held to the requirements and evaluated unless they received an exemption from the Executive Board prior to the event.These requirements are detailed in Article V under expectations for each category of membership. These requirements must be completed before the Membership Evaluation occurs. A secret Simple Majority vote with a quorum of two-thirds of all Active members is required for the evaluation to be valid. Neither absentee nor proxy votes are allowed. House may choose any of the Outcomes for each member.
\bsubsubsection{Outcomes}
Members who receive a pass have the option participate as an Active Member in the following the year and remain on the upcoming roster if applicable.
\\* \\*
If a member fails and has never passed an Membership Evaluation in the past, his membership will be revoked at the end of the year. If the member has previously passed an Membership Evaluation he will move to Alumni Membership status. In either case, the member forfeits his ability to be included on the following year’s roster.
\\* \\*
A member may be given a conditional to complete as a means of making up for missing requirements. A conditional may be proposed by any member present at the Evaluation and, if it is approved by the Evaluations Director, is then voted on by House. Each conditional consists of a set of additional requirements a deadline for completing them. When the deadline expires, the conditional will be brought before the Executive Board who will assess its completeness. A conditional member who does not meet these additional requirements will have failed the evaluation.
\bsubsection{Appeals Process}
If a member disagrees with the outcome of any evaluation, (e.g. is not asked to return to the House for the following year) and wishes to appeal the decision, he or she should immediately contact the Evaluations Director or the Chairman. At the following Executive Board meeting, the member will present his appeal to the Board. The Executive Board can uphold, overturn or conditionally overturn the Evaluation decision with a Simple Majority Vote as described in \ref{Simple Majority}.
\\* \\*
If the member is still unsatisfied after being heard by the Executive Board, the appeal may be brought to the attention of the Area Coordinator

\bsection{Expectations of House Members}
Each member is required to pay House dues as stated in \ref{Collection of House Dues}, attend all House Meetings, and attend at least twenty-five (25) of the permanent director meetings. Members are also required to actively participate in a majority of House activities.
\\* \\*
The member must participate significantly in at least one major House project during the academic year. The member is required to submit a description for this major project to the Executive Board for approval. As an option to this requirement, the member may instead assist on a large number of House activities and projects. However, it is to be understood in advance by the member that this option requires a great deal of participation throughout the year. This participation will be evaluated by the Executive Board.

\bsection{Housing Status}
All Active and Introductory Members have a housing status. This status indicates their right to priority housing on the floor. Any Alumni Member in good standing will maintain their previous On Floor status upon a return to Active membership.
\bsubsection{On-Floor Status}
Members with On-Floor status are eligible to live on the floor. To be granted On-Floor status, members may notify the Evaluations Director that they would like to come up for a vote. The Evaluations Director will then bring them up for vote at the next meeting.
\bsubsection{Off-Floor Status}
Members with Off-Floor status do not have the right to live on the floor. They still have access to all other privileges associated with their membership, and may still accumulate Housing Points.

\bsection{Housing Priority System}
The Housing Priority System is a means for determining the priority a House member has in a Housing issue such as Single Room Assignment or the Assignment of Available Housing. The member with top priority is the member with on-floor status and the most Housing Priority Points. Housing points are accumulated during the Housing Evaluation described in \ref{Housing Evaluation}.
\\* \\*
In the event of a tie, the members will be approached simultaneously and if they are unable to decide fairly between themselves, the assignment of priority will be made by random selection of the tied members.
\bsubsection{Single Room Assignments}
When a single room on the House becomes available it is offered to the member who carries the highest Housing Priority as defined in By-Law V, Section H. If that House member declines the option, it will be offered to the member with the next highest Housing Priority. This process continues until a member selects to move into the single room. Once in a single room, a House member retains the assignment until voluntarily giving it up. It should be noted that members selecting this option must agree to any additional charges applied by the Department of Residence Life for residing in a single room.
\bsubsection{Double Rooms as Single rooms}
During the third week of each semester, if there is no waiting list for residency on the House, any vacant rooms will be offered to House members as single rooms assignments according to the method as described in \ref{Single Room Assignments}. It should be noted that this does not mean members will be relocated into empty spaces so that the member with top priority is offered the single. This only applies if there is a totally vacant room.

% BY-LAW V - OPERATIONS OF THE OPERATIONAL COMMUNICATIONS DIRECTORSHIP
\bylaw{Operations of the Operational Communications Directorship}
The Operational Communications Directorship is responsible for overseeing the implementation of maintenance and upgrades to the House Computer Systems Network. It  is a self-governing committee making decisions by a simply majority vote. Membership is composed of all Root Type Persons.

\bsection{Selection of a Root Type Person}
\renewcommand{\theenumi}{\alph{enumi}} % For this section, we want items to use letters
\begin{enumerate}
	\item Nominations are taken from the Root Type Persons and Prior Root Type Persons meeting the selection criteria in \ref{Qualifications of a Root Type Person}.
	\item Each candidate is given a minimum of twenty-four hour period to accept or decline the nomination.
	\item A list of all nominees who have accepted is presented to the Executive Board for approval. This Executive Board Meeting is closed to the Executive Board Members, Root Type Persons, and House members with explicit invitation from the Executive Board.
	\item All current Executive Board Voting Members must be present during the discussion and voting period unless that member is a candidate for the office in discussion, in which case the member is absent and their vote is abstained. A simple majority Executive Board vote, as described in \ref{Simple Majority}, is taken to determine whether the nominations of the Root Type Persons are accepted.
\end{enumerate}

\bsection{Qualifications of a Root Type Person}
\renewcommand{\theenumi}{\alph{enumi}} % For this section, we want items to use letters
\begin{enumerate}
	\item Candidates must have been an On-Floor Active Member in good standing within the last twelve months, or been an Off-Floor Active Member and have lived on the floor within the last twelve months.
	\item Candidates may be granted an exemption by current Root Type Persons. Such an exemption may be withdrawn at any time by the current Root Type Persons.
	\item Prior Root Type Persons are those members who are no longer current Active Members and have not been granted an extension by the current Root Type Persons. Prior Root Type Persons are not guaranteed access to the current root passwords and other authentication tokens.
	\item The current Root Type Persons may from time to time draft rules and regulations specifying the rights and privileges of Prior Root Type Persons.
\end{enumerate}

\bsection{Creation of Accounts}
Root Type Persons have the authority to manage user accounts for House systems. Before members may receive an account they must:
\renewcommand{\theenumi}{\alph{enumi}} % For this section, we want items to use letters
\begin{enumerate}
	\item Sign the Code of Conduct sheets pertaining to the responsible utilization of Computer Science House and Rochester Institute of Technology facilities.
	\item Obtain 60\% (rounded up to the nearest whole person) of signatures in the Introductory Packet or successfully complete Introductory Evaluations.
\end{enumerate}

\end{document}